A preliminary constraint grammar for Russian

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
example ruls
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59567 72bbbca6-d526-0410-a7d9-f06f51895060

>>>>

\begin{figure*}
\centering
\begin{alltt}
{\small
\rus{
"<В>"                                                         "<работу>"
        "в" pr                                                     "работа" n f nn sg acc

"<ноябре>"                                                    "<Международный>"
        "ноябрь" n m nn sg prp                                     "международный" adj m an sg nom
                                                                   "международный" adj m nn sg acc
"<1994>"                                        
        "1994" num                                            "<трибунал>"
                                                                         "трибунал" n m nn sg acc
"<года>"                                                           "трибунал" n m nn sg nom
        "год" n m nn sg gen SELECT:r462                             
;        "год" n m nn pl nom fac SELECT:r462                   "<по>"
                                                                   "по" pr 
"<в>"                                                             
        "в" pr                                                "<Руанде>" 
                                                                   "Руанда" np al f nn sg prp 
"<Танзании>"                                                       "Руанда" np al f nn sg dat
        "Танзания" np al f nn pl acc                           
        "Танзания" np al f nn sg prp                          "<.>" 
;        "Танзания" np al f nn pl nom REMOVE:r424                   "." sent 
;        "Танзания" np al f nn sg dat REMOVE:r433 
;        "Танзания" np al f nn sg gen REMOVE:r433

"<начал>"
        "начало" n nt nn pl gen
        "начать" vblex perf tv past m sg
;        "начать" vblex perf iv past m sg REMOVE:r769
}}
\end{alltt}
  \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 
           года в Танзании начал работу Международный трибунал по Руанде.} ``The work of the International Tribunal 
           for Rwanda started in Tanzania in November 1994.'' The input ambiguity is 1.76 readings per word and 
           the output ambiguity is 1.38 readings per word. Recall is 1.0 and precision is 0.72. Figure~\ref{fig:rules} 
           shows the rules that fired for this example sentence.} 
  \label{fig:output}
\end{figure*}

\begin{figure*}

\begin{alltt}
{\small
### Safe

SELECT:r462 Gen IF (0 Year) (-1 Num LINK -1 Months LINK -1 Pr/V);
  # Select genitive reading of `\rus{года}' if there is a numeral immediately
  # to the left, before that there is a month and before that there is 
  # the preposition `\rus{в}'.

REMOVE:r424 Nom IF (-1C Pr) ;
  # Remove nominative case if there is a word which can only be a 
  # preposition immediately to the left.

REMOVE:r433 NGDAIP - Acc - Prp - Loc IF 
                (-1C* Pr/V OR Pr/Na
                  BARRIER (*) - Adv - Comp - DetIndecl - ModAcc - ModPrp);
  # Remove all cases apart from accusative, preposition and locative
  # if `\rus{в}' or `\rus{на}' are found to the left and are unambiguous. The barrier
  # is anything that cannot be found inside a noun phrase.

### Safe heuristic

REMOVE:r769 IV IF (0 TV OR IV) (1C Acc) (NOT 1 AccAdv);
  # Remove an intransitive reading of a verb if the next word can only
  # be accusative and is not in the set of nouns which can be used 
  # adverbially in accusative.
}
\end{alltt}
  \caption{Some example rules from the grammar.}
  \label{fig:rules}
\end{figure*}
